---
title: "#30DMC_12Nov_Time&Space"
author: "Kyri Janssen, Eugenio Muccio, Chiara Mazzarella"
date: "2024-11-11"
format: 
  html: default
  pdf: default
---

## 12 November - Time and space

"Map something where time matters. Visualize change over time—urban growth, migration, or environmental shifts. Show the relationship between time and geography."

### 1. Load packages

```{r}
library(tidyverse)
library(here)
library(sf)
library(ggplot2)
library(scales)  # Needed for the alpha() function
library(gridExtra)
library(magick)
library(grid)



```

## 12 November - Time and space

"Map something where time matters. Visualize change over time—urban growth, migration, or environmental shifts. Show the relationship between time and geography."

### 2. Import data & Rbanism logo

```{r}

banism_logo <- image_read('https://rbanism.org/assets/imgs/about/vi_l.jpg') # Download our logo}
layer_2023 <- read_sf(here('data', 'layer_2023_pc4_selection.shp'))
layer_2015 <- read_sf(here('data', 'CBS_PC4_2015_v2.shp'))


```

### 3. Data processing

```{r}
layer_2015 <- layer_2015%>%
  select(c(PC4_1, WONING))%>%
  mutate(WONING = ifelse(WONING==-99997, NA, WONING))%>%
  as.data.frame()%>%
  select(-geometry)

merge_2023_2015 <- merge(layer_2023,layer_2015, by.x='postcod', by.y="PC4_1", all.x=T)
merge_2023_2015<-merge_2023_2015%>%
  mutate(change_dwellings = ((antl_wn - WONING )/WONING) *100)%>%
 filter(!is.na(change_dwellings))



```

### 3. Visualize

```{r}
ggplot(data = merge_2023_2015) +
  geom_sf(aes(fill = change_dwellings), color = 'grey20') +  # Plotting without borders
  scale_fill_gradient2(
    low = "red",       # Color for negative values
    mid = "white",     # Color for values close to zero
    high = "blue",     # Color for positive values
    midpoint = 0,      # Set midpoint to zero
    name = "Percentage change"     # Legend title
  ) +
  labs(
    title = "Day 12. Time and space - Growth and loss of dwellings in Rotterdam, 2015–2023",
    subtitle = "Analyzing housing growth and decline across postcodes (PC4 level)",
    x = "", y = "",
    caption = "#30DayMapChallenge | Map by Kyri Janssen, Eugenio Muccio, Chiara Mazzarella, 2024 | Data: CBS"
  ) +
  theme_classic() +
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12)
  ) +
  # Adding a smaller subtitle with annotate
  annotate(
    "text", x = Inf, y = -Inf, label = "Note: harbor and non-residential areas are excluded from analysis",
    hjust = 1.1, vjust = -1.5, size = 3.5, color = "gray40", fontface = "italic"
  )

# Display the plot

grid.raster(rbanism_logo, x = 0.9, y=0.9, 
            width = unit(100, "points"))   # width determines the size of the logo

```

{r}

\`\`\`
